---
title: "CLA_exploration"
author: "Rebecca Bevans"
date: "October 1, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load libraries/data:
```{r echo=FALSE}
CLA<-read.csv("P:/Research/Thesis/PRP_thesis/Data/cla_2016_corrected.csv")
library(tidyverse)
library(broom)
library(ggplot2)
View(CLA)
```

Test CLA as function of diversity:
```{r echo=FALSE}
fit<-aov(cla_avg~diversity, data = CLA)
summary(fit)  
#No difference based on diversity

ggplot(CLA, aes(x=taxa, y=cla_avg))+ 
  geom_point(size = 2)
#this is confusing - why are ANGE and Solidago doubled????
#FIXED IT
```

Test CLA as function of location(latitutde)
```{r }
fitlat<-lm(CLA$cla_avg~ UTM_north, data=CLA)
summary(fitlat)
#Nothing is important here. 

#Try adding in random effects:
library(lmerTest)
fitlatlmer<-lmer(cla_avg ~ UTM_north + (1|block) + (1|level), data=CLA)
summary(fitlatlmer)
#Just not that much variation by location. That's cool. 
```

Test CLA as function of taxa:
```{r}
fitspecies<-aov(cla_avg~taxa, data=CLA)
summary(fitspecies)
#HUGE difference by taxa

#lmer
lmerfitspecies<-lmer(cla_avg~taxa + (1|block) + (1|level), data=CLA)
summary(lmerfitspecies)
#WHOA huge difference!!!

#Now let's make a boxplot of the differences:
#first do least-squares means tests
z<-lmerfitspecies
plot(z)
a<-lsmeans(z, c("taxa"))
a
az<-anova(z)
az
##I don't know why these values are so significant. 
##Why are group means 
#boxplot
dat=CLA
#View(dat)
dat$diversity=as.factor(dat$diversity)
dat$predaz=predict(z)

##now plot the predicted values based on the linear mixed-effects model
zplot<-ggplot(data=dat, aes(x=taxa, y=predaz, fill=taxa)) + scale_fill_grey(start=0.4, end=1.0) + geom_boxplot() + labs(x="Taxa", y="Chlorophyll by taxa + random effects of blockID")
zplot

```

Test CLA as function of treatment:
```{r}
fittrt<-aov(cla_avg~trt, data=CLA, na.rm=TRUE)
summary(fittrt)

#lmer
lmertrt<-lmer(cla_avg~trt + (1|block) + (1|level), data=CLA)
summary(lmertrt)
#Treatment is also significant. That's really cool.
```
There is no overall pattern based on latitude or diversity, but there is a difference based on species and by TREATMENT! 

Now we can dig into individual species to test responses to treatments.

Filter data to test solidago:
```{r echo=FALSE}
dfsoli<-filter(CLA, taxa=="solidago")

```
Now test solidago variation based on diversity level:
```{r echo=FALSE}
fit1<-aov(cla_avg~trt*diversity, data=dfsoli)
summary(fit1)
library(ggplot2)
```

Plotly plots:
```{r echo=FALSE}
library(plotly)
plot_ly(dfsoli, x= ~diversity, y = ~cla_avg, color= ~trt, type="box")%>%
  layout(boxmode="group")

plot_ly(dfsoli, x= ~diversity, y=~cla_avg, color=~trt, type="box")
        
```
Solidago chlorophyll varies significantly based on diversity level.

Let's plot it using a boxplot:
```{r echo=FALSE}
boxplot(cla_avg~trt:diversity, data=dfsoli)

```


Now test solidago variation based on latitude:
```{r echo=FALSE}
fit2<-lm(cla_avg~UTM_north, data=dfsoli)
summary(fit2)
```
Solidago chlorophyll content does not vary based on latitude.

Now test solidago variation based on treatment:
```{r echo=FALSE}
fit3<-aov(cla_avg~trt, data=dfsoli)
summary(fit3)
```
There is significant variation based on treatment. Let's check which treatment is most important:
```{r echo=FALSE}
boxplot(cla_avg~trt, data=dfsoli, ylab="Leaf Chlorophyll Content (nmol/cm^2)", xlab="Treatment", col=
          c("darkorange1", "firebrick3", "darkred", "dodgerblue3", "darkslategray4", "gray"), names= c("Biomass Removal", " ", "Biomass Removal/N Addition", "Control", "N Addition", "Rainout"))

library(plotly)
plot_ly(dfsoli, y=~cla_avg, color=~trt, type="box")
```

Now let's do this for Andropogon Gerardii:
```{r echo=FALSE}
dfange<-filter(CLA, taxa=="ange")

ange1<-aov(cla_avg~diversity, data=dfange)
summary(ange1)
#diversity is not important in ange chlorophyll

ange2<-lm(cla_avg~UTM_north, data=dfange)
summary(ange2)
#ange chlorophyll is not affected by the soil om gradient

ange3<-aov(cla_avg~trt, data=dfange)
summary(ange3)
#ange chlorophyll is significantly affected by treatment
#plot treatment effects:
boxplot(cla_avg~trt, data=dfange)

#why is my biomass factor duplicated?
duplicated(dfsoli)
```





