---
title: "Diversity assessment"
author: "Rebecca Bevans"
date: "November 8, 2016"
output: html_document
---

title: "prp_community_assessment"
author: "Rebecca Bevans"
date: "November 7, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
load data
```{r}
library(picante)
library(tidyverse)
meta_data<-read.csv("~/PRP_thesis/Data/meta_data.csv", row.names=2)
row.names(meta_data)
meta_data<-meta_data[c(-1)]
meta_data$unit<-paste(meta_data$block, meta_data$level)
#this isn't right, just manually load comm.csv from data folder
     #comm<-read.csv("~/PRP_thesis/Data/comm.csv", row.names=2)
class(comm)
dim(comm)
rownames(comm)
colnames(comm)
comm[1:5, 1:5]

all.equal(rownames(comm), rownames(meta_data))

```

Compare species richness between habitats:
```{r}
#species richness
specnumber(comm)
#check total abundance /sample
apply(comm, 1, sum)
comm.total<-decostand(comm, method="total")
apply(comm.total, 1, sum)
comm[1:5, 1:5]

colnames(comm)
```

```{r}

#fillCols1<-c(1='cornflowerblue', 2='indianred4')
#myPanel.divplot <- function(x=x, y=y, ...) {
#    fill <- fillCols1[intersect(levels(x), unique(x))] 
#    panel.bwplot(x=x, y=y, ..., groups=Col5, fill=fill)
#}

comm.noseed<-comm[c(-3, -7, -10, -11, -13, -31, -38, -43, -46, -47, -48, -56, -57, -58, -65, -70, -76, -77, -86, -91, -94, -95, -97, -99, -100, -105, -107, -108, -116, -127)]

boxplot(specnumber(comm.noseed)~meta$diversity, ylab = "# of species")

library(ggplot2)
ggplot(data=comm.noseed, aes(x=meta_data$diversity, y=specnumber(comm.noseed), fill=diversity)) + geom_boxplot() +
  stat_summary(fun.y=mean, colour="black", geom="point", 
               shape=18, size=3,show.legend = FALSE) + 
  geom_text(data = means, aes(label = bare, y = bare + 0.08))

mod<-aov(specnumber(comm.noseed)~meta_data$diversity, ylab = "# of species")
summary(mod)
pander(mod)

boxplot(specnumber(comm.noseed)~meta_data$year, ylab = "# of species")

boxplot(specnumber(comm.noseed)~meta_data$diversity:meta_data$treat)

mod1<-aov(specnumber(comm.noseed)~meta$treat + meta$replicate + meta$treat:meta$replicate)
summary(mod1)

div.trt<-aov(specnumber(comm.noseed)~meta_data$diversity + meta_data$treat + meta_data$diversity:meta_data$treat)
summary(div.trt)
library(pander)
library(mgcv)
bare.div.trt<-pander(div.trt)
pdf(bare.div.trt, file = "~/PRP_thesis/Graphs/bare.div.trt.pdf")

##create the color scheme for the bwplots
fillCols<-c(biomass='cornflowerblue', control='indianred4', nitrogen='olivedrab', biomass_nitrogen='coral1', rainout='darkgoldenrod1', no='burlywood')
myPanel.bwplot <- function(x=x, y=y, ...) {
    fill <- fillCols[intersect(levels(x), unique(x))] 
    panel.bwplot(x=x, y=y, ..., groups=Col5, fill=fill)
}

#this doesn't work. #fill<-c(2015.1 = 'brickred', 2015.2 ='goldenrod1', 2016.1='cornflowerblue', 2016.2 = 'cyan')

boxplot(specnumber(comm)~meta_data$bare + meta_data$diversity)

boxplot(specnumber(comm.noseed)~meta_data$year:meta_data$diversity, notch=FALSE, col = (c('indianred4','goldenrod1','cornflowerblue','cyan')), main="Species Richness by Year and Diversity")

#don't know why but this ins't working. Manually created image
  #jpeg(richbyyr, file="~/PRP_thesis/Graphs/richbyyr.jpeg")

comm.hd<-subset(comm.noseed, meta_data$diversity == "2")
  
bwplot(specnumber(comm.hd) ~ meta_data$treat | factor(meta_data$year), comm.hd)


bwplot(specnumber(comm.hd)~meta_data$treat | factor(meta_data$unit), data=comm.noseed, panel=myPanel.bwplot)

#species by bareground
divv<-manova(specnumber(comm.hd) + meta_data$bareground ~ ))
```

Statistical test of difference based on diversity and year:
```{r}
library(pander)
#welch's t-test:
div<-t.test(specnumber(comm)~meta_data$diversity)
pander(div)
yr<-t.test(specnumber(comm)~meta_data$year)
pander(yr)
```
Both year and diversity plantings have significant impact on species richness at the level of the individual quadrat. Now I want to see if these differences hold at the level of treatments.

We have our species comparisons by quadrat, but these are not individual observations. Need to average to the transect level. I think the easiest way to do this will be to subset my larger dataset based on the block, level, and transects:
```{r}
#load subsetted data from my bareground rmarkdown
subplot.comm <- read.csv("~/PRP_thesis/Data/subplot.comm.csv", row.names=1)
subplot <- read.csv("~/PRP_thesis/Data/subplot.csv", row.names=1)
all.equal(rownames(subplot.comm), rownames(subplot))
bwplot(specnumber(subplot.comm)~subplot$treat | factor(subplot$replicate), subplot.comm, panel=myPanel.bwplot)
mod<-aov(specnumber(subplot.comm)~subplot$treat + subplot$replicate + subplot$treat:subplot$replicate)
summary(mod)

##now wholeplot data:
wholeplotcomm <- read.csv("~/PRP_thesis/Data/wholeplot.comm", row.names=1)
wholeplot <- read.csv("~/PRP_thesis/Data/wholeplot.csv", row.names=1)
all.equal(rownames(wholeplotcomm), rownames(wholeplot))

#average by transect within each wholeplot
bwplot(specnumber(wholeplotcomm)~wholeplot$transect | factor(wholeplot$replicate), wholeplotcomm, panel=myPanel.bwplot)

#average by wholeplot
bwplot(specnumber(wholeplotcomm)~wholeplot$treat:wholeplot$replicate, data=wholeplotcomm, notch=TRUE, col=meta_data$block, main = "Species richness by wholeplot unit")

#anova to test if statistical differences exist
wholemod<-aov(specnumber(wholeplotcomm)~ wholeplot$replicate)
summary(wholemod)

library(mgcv)
wholemod2<-gam(specnumber(wholeplotcomm)~wholeplot$year + wholeplot$replicate + wholeplot$year:wholeplot$replicate)
summary(wholemod2)

bwplot(specnumber(wholeplotcomm)~wholeplot$replicate | factor(wholeplot$year), wholeplotcomm)
```

This is copied from a tutorial on analyzing community data:
```{r}
#Species-area and species-individual curves:
spa<-specaccum(comm)
plot(spa)
plot(spa, ci.type="poly", col="blue", lwd=2, ci.lty=0, ci.col="lightblue")

#Now try the same thing doing individuals rather than sites
#spi<-specaccum(comm, method="rarefaction")
#plot(spi)
##This won't work with my data unless I transpose all of the cover data to 1s and 0s##

#Ordination
#Distance measures:
#Bray-Curtis index = percent similarity including species abundances. #Sorenson's index = percent similarity based on presence-absence data.

#calc. Bray-Curtis distance among samples
comm.bc.dist<-vegdist(comm, method = "bray")
#cluster communities using average-linkage algorithm
comm.bc.clust<-hclust(comm.bc.dist, method = "average")
#plot cluster diagram
plot(comm.bc.clust, ylab = "Bray-Curtis dissimilarity")
#create a map of the plots in community composition space using Non-metric #Multidimensional scaling.

bci.mds<-metaMDS(comm, distance = "bray", k=2, trymax = 20, autotransform = TRUE, noshare = 0.1, expand = TRUE, trace = 1, plot = FALSE)

mds.fig<-ordiplot(bci.mds, type = "none")
##don't use this one:
        #plot(bci.mds, choices = c(1,2), type="n")
points(bci.mds, display = c("sites"))

text(bci.mds, display = c("sites"))


##Now I'm going to try to visualize these communities by site and treatment
#plot just the samples, colour by habitat, pch=19 means plot a circle:
points(mds.fig, "sites", pch = 19, col = "blue", 
       select = cover_data$diversity=="1")
points(mds.fig, "sites", pch=19, col="goldenrod1",
       select=cover_data$diversity=="2")
#add confidence ellipses around habitat types
ordiellipse(bci.mds, cover_data$diversity, conf = 0.95, label=TRUE)

#overlay the cluster results calculated earlier
ordicluster(bci.mds, comm.bc.clust, col="gray")
```

stressplot to test goodness of fit:
```{r}
stressplot(bci.mds)
```

Now I'm going to do the whole thing again dividing up by assessment scale and year:
```{r}
#calc. Bray-Curtis distance among samples
wholecomm.bc.dist<-vegdist(wholeplotcomm, method = "bray")
#cluster communities using average-linkage algorithm
wholecomm.bc.clust<-hclust(wholecomm.bc.dist, method = "average")
#plot cluster diagram
plot(wholecomm.bc.clust, ylab = "Bray-Curtis dissimilarity")
#create a map of the plots in community composition space using Non-metric #Multidimensional scaling.

wbci.mds<-metaMDS(wholeplotcomm, distance = "bray", k=2, trymax = 20, autotransform = TRUE, noshare = 0.1, expand = TRUE, trace = 1, plot = FALSE)

wmds.fig<-ordiplot(wbci.mds, type = "none")
##or:
#plot(wbci.mds, choices = c(1,2), type="n")

points(wbci.mds, display = c("sites"))

text(wbci.mds, display = c("sites"))


##Now I'm going to try to visualize these communities by site and treatment
#plot just the samples, colour by habitat, pch=19 means plot a circle:
points(wmds.fig, "sites", pch = 19, col = "aquamarine", 
       select = wholeplot$diversity=="1" & wholeplot$year=="2015")
points(wmds.fig, "sites", pch = 19, col = "blue", 
      select = wholeplot$diversity=="2" & wholeplot$year=="2015")    
points(wmds.fig, "sites", pch=19, col= "goldenrod1",
       select = wholeplot$diversity=="1" & wholeplot$year=="2016")
points(wmds.fig, "sites", pch=19, col="olivedrab", 
      select = wholeplot$diversity=="2" & wholeplot$year=="2016")

#try it a different way, organizing by experimental replicate
points(wmds.fig, "sites", pch = 19, col = "goldenrod2", 
       select = wholeplot$replicate=="a 1")
points(wmds.fig, "sites", pch = 19, col = "orange", 
      select = wholeplot$replicate=="a 3")    
points(wmds.fig, "sites", pch=19, col= "olivedrab",
       select = wholeplot$replicate=="a 4")
points(wmds.fig, "sites", pch=19, col="darkslategray1", 
      select = wholeplot$replicate=="b 2")
points(wmds.fig, "sites", pch=19, col="coral1", 
      select = wholeplot$replicate=="b 4")      
points(wmds.fig, "sites", pch=19, col="cyan", 
      select = wholeplot$replicate=="c 1")
points(wmds.fig, "sites", pch=19, col="indianred4", 
      select = wholeplot$replicate=="c 2")
points(wmds.fig, "sites", pch=19, col="seashell2", 
      select = wholeplot$replicate=="c 4")      

      
#add confidence ellipses around habitat types
ordiellipse(wbci.mds, wholeplot$diversity, conf = 0.95, label=TRUE)
ordiellipse(wbci.mds, wholeplot$year, conf=0.95, label=TRUE)
#overlay the cluster results calculated earlier
ordicluster(wbci.mds, comm.bc.clust, col="gray")
```

stressplot to test goodness of fit:
```{r}
stressplot(wbci.mds)
```

Now at the subplot scale
```{r}
#calc. Bray-Curtis distance among samples
subcomm.bc.dist<-vegdist(subplot.comm, method = "bray")
#cluster communities using average-linkage algorithm
subcomm.bc.clust<-hclust(subcomm.bc.dist, method = "average")
#plot cluster diagram
plot(subcomm.bc.clust, ylab = "Bray-Curtis dissimilarity")
#create a map of the plots in community composition space using Non-metric #Multidimensional scaling.

sbci.mds<-metaMDS(subplot.comm, distance = "bray", k=2, trymax = 20, autotransform = TRUE, noshare = 0.1, expand = TRUE, trace = 1, plot = FALSE)

smds.fig<-ordiplot(sbci.mds, type = "none")
##or:
#plot(sbci.mds, choices = c(1,2), type="n")

points(sbci.mds, display = c("sites"))

text(sbci.mds, display = c("sites"))


##Now I'm going to try to visualize these communities by site and treatment
#plot just the samples, colour by habitat, pch=19 means plot a circle:
points(smds.fig, "sites", pch = 19, col = "aquamarine", 
       select = subplot$diversity=="1" & subplot$year=="2015")
points(smds.fig, "sites", pch = 19, col = "blue", 
      select = subplot$diversity=="2" & subplot$year=="2015")    
points(smds.fig, "sites", pch=19, col= "goldenrod1",
       select = subplot$diversity=="1" & subplot$year=="2016")
points(smds.fig, "sites", pch=19, col="olivedrab", 
      select = subplot$diversity=="2" & subplot$year=="2016")

#add confidence ellipses around habitat types
ordiellipse(sbci.mds, subplot$diversity, conf = 0.95, label=TRUE)
ordiellipse(sbci.mds, subplot$year, conf=0.95, label=TRUE)


#try it a different way, organizing by experimental replicate

smds.fig<-ordiplot(sbci.mds, type = "none")
points(sbci.mds, display = c("sites"))

subplot$replicate<-paste(subplot$block, subplot$level)

points(smds.fig, "sites", pch = 19, col = "goldenrod2", 
       select = subplot$replicate=="a 1")
points(smds.fig, "sites", pch = 19, col = "orange", 
      select = subplot$replicate=="a 3")    
points(smds.fig, "sites", pch=19, col= "olivedrab",
       select = subplot$replicate=="a 4")
points(smds.fig, "sites", pch=19, col="darkslategray1", 
      select = subplot$replicate=="b 2")
points(smds.fig, "sites", pch=19, col="coral1", 
      select = subplot$replicate=="b 4")      
points(smds.fig, "sites", pch=19, col="cyan", 
      select = subplot$replicate=="c 1")
points(smds.fig, "sites", pch=19, col="indianred4", 
      select = subplot$replicate=="c 2")
points(smds.fig, "sites", pch=19, col="seashell2", 
      select = subplot$replicate=="c 4")      

      
#add confidence ellipses around habitat types
ordiellipse(sbci.mds, subplot$diversity, conf = 0.95, label=TRUE)
ordiellipse(sbci.mds, subplot$year, conf=0.95, label=TRUE)
#overlay the cluster results calculated earlier
ordicluster(wbci.mds, comm.bc.clust, col="gray")
```

stressplot to test goodness of fit:
```{r}
stressplot(wbci.mds)
```

Large overlaps in the species diversity at the quadrat level among diversity treatments and year of experiment. It seems that time may be the biggest variable determining diferences than anything else.


How many plots does each species occur in?
```{r}

library(labdsv)
comm.cover<-comm
x<-c(1, 2, 3, 4, 5, 6)
y<- c(3, 15, 37.5, 62.5, 85, 97.5)
comm.cover<-vegtrans(comm.cover, x, y)


library(tidyverse)

spc_pres<-apply(comm.cover > 0,2, sum)
plot(sort(spc_pres))
as.matrix(spc_pres)
plot(sort(spc_pres), log='y')    

seq(1,138)[sort(spc_pres)==10]
spc_pres[spec_pres==10]
spcies<-sort(spc_pres)
library(pander)
pander(species)
#pdf(species, file = "~/PRP_thesis/Graphs/species.pdf")
hist(spc_pres)
hist(log(spec_pres))
```

What is the mean cover of each species when it occurs(ignoring zeroes where it is absent)?
```{r}
tmp<-apply(comm.cover, 2, sum)
spc_mean<-tmp/spc_pres
plot(sort(spc_mean))
summary(spc_mean)
as.matrix(spc_mean)

```

Is the mean abundance of species correlated with the number of plots they occur in?
```{r}
plot(spc_pres, spc_mean)
identify(spc_pres, spc_mean)
identify(spc_pres, spc_mean, names(comm.cover))

```

