---
title: "Diversity_practice_VEGAN"
author: "Rebecca Bevans"
date: "November 22, 2016"
output: html_document
---

---
title: "Vegetation_diversity_practice"
author: "Rebecca Bevans"
date: "November 22, 2016"
output: html_document
---
load libraries and data:
```{r}
library(vegan)
library(MASS)
data(BCI)
dim(BCI)
BCI[1:10, 20:25]
BCI[1:5, 70:75]
BCI$Faramea.occidentalis
sum(BCI$Faramea.occidentalis)
sum(BCI$Poulsenia.armata)
```

Shannon Diversity Indices:
```{r}
#get diversity for each plot
diversity(BCI, index = "shannon")
```

Use these samples to get average diversity for a hectare on Barro Colorado Island:
```{r }
plots<-diversity(BCI, index = "shannon")
summary(plots)
median(plots)
mean(plots)
```

Fisher's Diversity:
```{r}
fish.a<-fisher.alpha(BCI, MARGIN = 1)
fish.a
```

Use Fisher's to get average diversity:
```{r}
bcitot<- apply(BCI, 2, sum)
bcitot.a<-fisher.alpha(bcitot, MARGIN=1)
bcitot
bcitot.a
```

Calculate Fisher's alpha for increasing numbers of hectares combined:
```{r}
x<-1:50
a<-NULL
for(i in x){
b<-apply(BCI[1:i,],2, sum)
c<-fisher.alpha(b, MARGIN=1)
a<-c(a,c) 
}
plot(x,a)
```

Species richness:
```{r}
rar<-rarefy(BCI, 20)
rarsum<-rarefy(bcitot, 20, MARGIN=2)
rar
rarsum
```

Species-area and species-individual curves:
```{r}
spa<-specaccum(BCI)
plot(spa)
plot(spa, ci.type="poly", col="blue", lwd=2, ci.lty=0, ci.col="lightblue")

#Now try the same thing doing individuals rather than sites
spi<-specaccum(BCI, method="rarefaction")
plot(spi)
```

Ordination
Distance measures:
Bray-Curtis index = percent similarity including species abundances. Sorenson's index = percent similarity based on presence-absence data.
```{r}
bc<-vegdist(BCI, method = "bray", binary=FALSE)
bc

```

create a map of the plots in community composition space using Non-metric Multidimensional scaling.
```{r}
bci.mds<-metaMDS(BCI, distance = "bray", k=2, trymax = 20, autotransform = TRUE, noshare = 0.1, expand = TRUE, trace = 1, plot = FALSE)

plot(bci.mds, choices = c(1,2), type="n")
points(bci.mds, display = c("sites"))
text(bci.mds, display = c("sites"))

```

