---
title: "Bevans_Final_Project"
author: "Rebecca Bevans"
date: "December 13, 2016"
output: word_document
---

This code can be used to reproduce the analyses and figures created for Rebecca Bevans' Ecostats final project (fall 2016): "Biodiversity, Resource Availability, and Invasion Density" in R studio. 

Load data and packages:
```{r}
library(tidyverse)
library(nlme)
library(picante)
library(AICcmodavg)
library(lme4)
library(labdsv)
invsoil<-read.csv("~/PRP_thesis/Data/invsoil.csv", header=TRUE, row.names=1)
```

Test correlations among variables:
```{r}
library(GGally)
ggpairs(invsoil)
#77.5% correlation between individual Omloi and avg.c
#66.7% correlation between individual kclN and avg.n
#9.5% correlation between avg.c and avg.n
#-19.9%correlation between avg.n and diversity
#5.07% corr between avg.c and diversity
#some significant variation between plot position and soilOM and N; this should be accounted for by using block as random effects variable
```

Build model sets:
```{r}
#Global model using indexes
all<-lmer(all.index ~ avg.c + avg.n + diversity + avg.c*diversity + avg.n*diversity + avg.c*avg.n + (1|rep), data=invsoil, REML=FALSE)
summary(all)
plot(resid(all))
drop1(all, test="Chisq")

all1.1<-lmer(all.index ~ avg.c + avg.n + diversity + avg.c*diversity + avg.n*diversity + (1|rep), data=invsoil, REML=FALSE)
summary(all1.1)
drop1(all1.1, test="Chisq")

all1.2<-lmer(all.index ~ avg.c + diversity + avg.c:diversity + (1|rep), data=invsoil, REML=FALSE)
summary(all1.2)


all1<-lmer(all.index ~ avg.c + avg.n + diversity + avg.n*diversity + avg.c*avg.n + (1|invsoil$rep), data=invsoil, REML=FALSE)
summary(all1)
plot(resid(all1))
drop1(all1, test="Chisq")
anova(all1, all1.1, test="F")

all2<-lmer(all.index ~ avg.c + avg.n + diversity + avg.n*diversity + (1|invsoil$rep), data=invsoil, REML=FALSE)
summary(all2)
plot(resid(all2))
anova(all1, all2, test="F")
anova(all, all2, test="F")
drop1(all2, test="Chisq")

all3<-lmer(all.index ~ avg.n + avg.n*diversity + (1|rep), data=invsoil, REML=FALSE)
summary(all3)
anova(all1.2, all3, test="F")
drop1(all3, test="Chisq")
#The all3 model is the best for all invaders overall. Though it is hardly predictive. It's pretty bad. 

all4<-lmer(all.index ~ (1|rep), data=invsoil, REML=FALSE)
summary(all4)
anova(all1.2, all4, test="F")
anova(all1.2, all4)
#Test for woody invaders alone:
lmwoody<-lmer(woody.index ~ avg.c + avg.n + diversity + avg.c*diversity + avg.n*diversity + avg.c*avg.n + (1|block/level), data=invsoil, REML=FALSE)
summary(lmwoody)
plot(resid(lmwoody))

drop1(lmwoody)

lmwoody2<-lmer(woody.index ~ avg.c + avg.n + diversity + avg.c*diversity + avg.n*diversity + (1|block/level), data=invsoil, REML=FALSE)
summary(lmwoody2)
plot(resid(lmwoody2))
drop1(lmwoody2)

anova(lmwoody, lmwoody2, test="F")

lmwoody3<-lmer(woody.index ~ avg.c + avg.n + diversity + avg.c*diversity + (1|block/level), data=invsoil, REML=FALSE)
  
summary(lmwoody3)
drop1(lmwoody3)
plot(resid(lmwoody3))
anova(lmwoody2, lmwoody3, test="F")
#Best model is lmwoody3; difference in model weights between lmwoody3 and lmwoody4 is not statistically significant.

lmwoody4<-lmer(woody.index ~ avg.c + avg.n + diversity + (1|invsoil$rep), data=invsoil, REML=FALSE)

summary(lmwoody4)

anova(lmwoody3, lmwoody4, test="F")


#Test for thistles only:
thistle<-lmer(cirsium.index ~ avg.c + avg.n + diversity + avg.c*diversity + avg.n*diversity + avg.c*avg.n + (1|invsoil$rep), data=invsoil, REML=FALSE)
summary(thistle)
plot(resid(thistle))
drop1(thistle, test="Chisq")

thistle2<-lmer(cirsium.index ~ avg.c + avg.n + diversity + avg.c*diversity + avg.n*diversity + (1|invsoil$rep), data=invsoil, REML=FALSE)
summary(thistle2)
plot(resid(thistle2))
drop1(thistle2, test="Chisq")

thistle3<-lmer(cirsium.index ~ avg.c + avg.n + diversity + avg.c*diversity + (1|invsoil$rep), data=invsoil, REML=FALSE)
summary(thistle3)
plot(resid(thistle3))
drop1(thistle3, test="Chisq")
#thistle3 is the best fitting model

```


Graphics:
```{r}
library(ggplot2)
library(gridExtra)
#correlation between OM and invasive density
#all invaders
oplot<-ggplot(invsoil, aes(x=avg.c, y=all.index))
oplot + geom_point() + geom_smooth()
o<-oplot + geom_point() + geom_smooth()

#woody invaders only
woplot<-ggplot(invsoil, aes(x=avg.c, y=woody.index))
woplot +geom_point() + geom_smooth()
wo<-woplot +geom_point() + geom_smooth()

#thistles
toplot<-ggplot(invsoil, aes(x=avg.c, y=cirsium.index))
toplot + geom_point() + geom_smooth()
to<-toplot+geom_point() +geom_smooth()
#all on same page:
grid.arrange(o, wo, to, ncol=1)


#correlation between N and invasive density
#all invaders
nplot<-ggplot(invsoil, aes(x=avg.n, y=all.index))
nplot + geom_point() + geom_smooth()
n<-nplot + geom_point() + geom_smooth()

#woody invaders only
wnplot<-ggplot(invsoil, aes(x=avg.n, y=woody.index))
wnplot +geom_point() + geom_smooth()
w<-wnplot +geom_point() + geom_smooth()

#thistles
tnplot<-ggplot(invsoil, aes(x=avg.n, y=cirsium.index))
tnplot + geom_point() + geom_smooth()
t<-tnplot+geom_point() +geom_smooth()
#all on same page:
grid.arrange(n, w, t, ncol=1)

###
#correlation (bwplot) between diversity and invasive density
dplot<-ggplot(invsoil, aes(x=factor(diversity), y=all.index))
dplot + geom_boxplot()
d<-dplot + geom_boxplot()

wdplot<-ggplot(invsoil, aes(x=factor(diversity), y=woody.index))
wdplot + geom_boxplot()
wd<-wdplot + geom_boxplot()

tdplot<-ggplot(invsoil, aes(x=factor(diversity), y=cirsium.index))
tdplot + geom_boxplot()
td<-tdplot + geom_boxplot()

#all in same image
grid.arrange(d, wd, td, ncol=1)
```

Linear interaction graphics:
```{r}
#graphics theme - APA style
apatheme=theme_bw()+
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.border=element_blank(),
        axis.line=element_line())
        

#Average Nxdiversity effect on total invasion:
dat=invsoil
dat$diversity=as.factor(dat$diversity)
nmod<-glm(all.index~avg.n*diversity, data=dat)
dat$predicted<-predict(nmod)
#all3 lmer: all.index ~ avg.n + avg.n*diversity + (1|rep)
#make diversity a discrete variable for plotting



#plot scatterplot with fit lines
p=ggplot(dat, aes(x=avg.n, y=all.index, shape=diversity))+
geom_point()+
scale_shape_manual(values=c(1,0.2,16), name='Diversity', labels=c('LD','MD','HD'))+
geom_line(aes(x = avg.n, y = predicted, linetype=diversity))+
scale_linetype_discrete(name='Diversity', labels=c('LD','MD', 'HD'))+
labs(x = 'Average Soil N', y = 'Total Invasion Density')+apatheme
p


#do it with just thistles
dat=invsoil
dat$diversity<-as.factor(dat$diversity)
nmod2<-glm(cirsium.index~avg.n*diversity, data=dat)
dat$pred<-predict(nmod2)
#plot scatterplot with fit lines
p1.1=ggplot(dat, aes(x=avg.n, y=cirsium.index, shape=diversity))+
geom_point()+
scale_shape_manual(values=c(1,0.2,16), name='Diversity', labels=c('LD','MD','HD'))+
geom_line(aes(x = avg.n, y = pred, linetype=diversity))+
scale_linetype_discrete(name='Diversity', labels=c('LD','MD', 'HD'))+
labs(x = 'Average Soil N', y = 'Thistle Invasion Density')+apatheme
p1.1


#do it with just woody
dat=invsoil
dat$diversity<-as.factor(dat$diversity)
nmod3<-glm(woody.index~avg.n*diversity, data=dat)
dat$pred<-predict(nmod3)
#plot scatterplot with fit lines
p1.2=ggplot(dat, aes(x=avg.n, y=woody.index, shape=diversity))+
geom_point()+
scale_shape_manual(values=c(1,0.2,16), name='Diversity', labels=c('LD','MD','HD'))+
geom_line(aes(x = avg.n, y = pred, linetype=diversity))+
scale_linetype_discrete(name='Diversity', labels=c('LD','MD', 'HD'))+
labs(x = 'Average Soil N', y = 'Woody Invasion Density')+apatheme
p1.2

#Average C*diversity effect on thistle invasion:
dat=invsoil
predt<-glm(cirsium.index~avg.c*diversity, data=dat)
dat$predt<-predict(predt)
#all3 lmer: all.index ~ avg.n + avg.n*diversity + (1|rep)
#make diversity a discrete variable for plotting
dat$diversity<-as.factor(dat$diversity)

#plot scatterplot with fit lines
p=ggplot(dat, aes(x=avg.c, y=cirsium.index, shape=diversity))+
geom_point()+
scale_shape_manual(values=c(1,0.2,16), name='Diversity', labels=c('LD','MD','HD'))+
geom_line(aes(x = avg.c, y = predt, linetype=diversity))+
scale_linetype_discrete(name='Diversity', labels=c('LD','MD', 'HD'))+
labs(x = 'Average Soil OM', y = 'Thistle Invasion Density')+ apatheme
p

```


Barplot graphics code:
```{r}
library(psych)
#p1 data
dat<-invsoil
dat$diversity=as.factor(dat$diversity)
dat$avg.c=as.factor(dat$avg.c)
dat2<-describeBy(dat$woody.index, list(dat$diversity), mat=TRUE, digits=3)
names(dat2)[names(dat2)=='group1']= 'Diversity'


dat2$se=dat2$sd/sqrt(dat2$n)

#plot layout
limits = aes(ymax = mean + (1.96*se), ymin=mean - (1.96*se))
 
dodge = position_dodge(width=0.9)
 
apatheme=theme_bw()+
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.border=element_blank(),
        axis.line=element_line())

give.n<-function(x){
  return(c(y=mean(x), label=length(x)))
}

#plot the interaction
p1<-ggplot(dat2, aes(x=Diversity, y=mean, fill=Diversity))+
  geom_bar(stat='identity', position=dodge)+
  geom_errorbar(limits, position=dodge, width=0.25)+
  ylab('Mean Shrub Density')+
  scale_fill_grey()
p1

#thistle invasion by diversity
dat<-invsoil
dat$diversity=as.factor(dat$diversity)
dat$cirsium.index=as.numeric(dat$cirsium.index)
dat3<-describeBy(dat$cirsium.index, list(dat$diversity), mat=TRUE, digits=3)
names(dat3)[names(dat3)=='group1']= 'Diversity'

dat3$se=dat3$sd/sqrt(dat3$n)

p2<-ggplot(dat3, aes(x=Diversity, y=mean, fill=Diversity))+
geom_bar(stat='identity', position=dodge)+
  geom_errorbar(limits, position=dodge, width=0.25)+
  ylab('Mean Thistle Density')+
  scale_fill_grey() 
p2

#all invasion by diversity
dat<-invsoil
dat$diversity=as.factor(dat$diversity)
dat4<-describeBy(dat$all.index, list(dat$diversity), mat=TRUE, digits=3)
names(dat4)[names(dat4)=='group1']= 'Diversity'

dat4$se=dat4$sd/sqrt(dat4$n)

p3<-ggplot(dat4, aes(x=Diversity, y=mean, fill=Diversity))+
geom_bar(stat='identity', position=dodge)+
  geom_errorbar(limits, position=dodge, width=0.25)+
  ylab('Mean Invader Density')+
  scale_fill_grey() 
p3
#p1 and p2 on same plot:
grid.arrange(p1, p2, p3, ncol=3)

#tdplot<-ggplot(invsoil, aes(x=factor(diversity), #y=cirsium.index))
#tdplot + geom_boxplot()
#td<-tdplot + geom_boxplot()
```


ANOVAs for combined fixed and random factors:
```{r}
library(nlme)
library(lmerTest)

#for cirsium species
z <- lme(cirsium.index ~ diversity*avg.c, random = ~ 1|rep, data = invsoil) # nlme package
z <- lmer(cirsium.index ~ diversity*avg.c + (1|rep), data = invsoil) 
plot(z)
dat=invsoil

#model-based fitted means for treatments:
lsmeans(z, c("diversity", "avg.c"))
lsmeans(z, c("diversity"))
az<-anova(z)
az
#boxplot
dat=invsoil
dat$diversity=as.factor(dat$diversity)
dat$predaz=predict(z)  
zplot<-ggplot(data=dat, aes(x=diversity, y=predaz, fill=diversity)) + scale_fill_grey(start=0.4, end=1.0) + geom_boxplot() + labs(x="Diversity", y="Thistle Invasion by diversity*soil OM + BlockID.rand.eff.")
zplot


###for woody invaders
z2 <- lme(woody.index ~ diversity*avg.c, random = ~ 1|rep, data = invsoil) # nlme package
z2 <- lmer(woody.index ~ diversity*avg.c + (1|rep), data = invsoil) 
plot(z2)
dat=invsoil
dat$diversity=as.factor(dat$diversity)
#model-based fitted means for treatments:
lsmeans(z2, c("diversity", "avg.c"))
lsmeans(z2, c("diversity"))
az2<-anova(z2)
az2

#boxplot
dat$predaz2=predict(z2)  
zplot2<-ggplot(data=dat, aes(x=diversity, y=predaz2, fill=diversity)) + scale_fill_grey(start=0.4, end=1.0) + geom_boxplot() + labs(x="Diversity", y="Woody Invasion by diversity*soil OM + BlockID.rand.eff.")
zplot2


###for all invaders
z1 <- lme(all.index ~ diversity*avg.c, random = ~ 1|rep, data = invsoil) # nlme package
z1 <- lmer(all.index ~ diversity*avg.c + (1|rep), data = invsoil) 
plot(z)
dat=invsoil

#model-based fitted means for treatments:
lsmeans(z, c("diversity", "avg.c"))
lsmeans(z, c("diversity"))
az1<-anova(z1)
az1

#boxplot
dat$diversity=as.factor(dat$diversity)
dat$predaz1=predict(z1)  
zplot1<-ggplot(data=dat, aes(x=diversity, y=predaz1, fill=diversity)) + scale_fill_grey(start=0.4, end=1.0) + geom_boxplot() + labs(x="Diversity", y="Total Invasion by diversity*soil OM + BlockID.rand.eff.")
zplot1


#plot all 3 together:
grid.arrange(zplot1, zplot, zplot2, ncol=3)
```