---
title: "SLA_analysis"
author: "Becca"
date: "January 9, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load data and arrange weight and area datasets to match. 
```{r echo=FALSE}
#read in sla weight file and add column "unit"

#sla_weight<-read.csv("P:/Research/Thesis/PRP_thesis/Data/sla_2016_weight.csv", header=TRUE)
#sla_weight<-sla_weight[c(-11,-12,-13,-14,-15,-16)]
#View(sla_weight)
#sla_weight$unit<-paste(sla_weight$block, sla_weight$level, sla_weight$transect, sla_weight$quadrat, sla_weight$taxa, sla_weight$replicate)
#write.csv(sla_weight, file="P:/Research/Thesis/PRP_thesis/Data/sla_weight.csv")

#read in sla area file and add column "unit"
#sla_area<-read.csv("P:/Research/Thesis/PRP_thesis/Data/SLA_2016_area.csv", header=TRUE)
#View(sla_area)
#sla_area<-sla_area[c(-11)]
#sla_area$unit<-paste(sla_area$block, sla_area$level, sla_area$transect, sla_area$quadrat, sla_area$taxa, sla_area$replicate)
#write.csv(sla_area, file = "P:/Research/Thesis/PRP_thesis/Data/sla_area.csv")

#sort to match
#sla<-merge(sla_area, sla_weight, by="unit")
#View(sla)
#write.csv(sla, file="P:/Research/Thesis/PRP_thesis/Data/sla.csv")

sla<-read.csv("P:/Research/Thesis/PRP_thesis/Data/sla.csv", header=TRUE)
View(sla)
```
Yay! All organized!

2.Load libraries
```{r echo=FALSE}

library(ggplot2)
library(tidyverse)
library(picante)
library(Ggally)
library(lattice)
#install.packages('data.table') #may need to be run if you don't have the package
library(data.table)
library(lmerTest)
library(MASS)
#install.packages('pander')
library(pander)
```

3. Create SLA metric (area/weight, for unit sLA)
```{r echo=FALSE}
#make area and weight numeric
#sla$area.cm = as.numeric(sla$area.cm)
#sla$weight.0.00g = as.numeric(sla$weight.0.00g)
#sla<-transform(sla, leaf.sla=sla$area.cm/sla$weight.0.00g)
#View(sla)
#sla$leaf.sla<-as.numeric(sla$leaf.sla)
#write.csv(sla, file = "P:/Research/Thesis/PRP_thesis/Data/sla.csv")
```

3. Screen data for differences by species, treatment, diversity, and look for outliers or non-heterogenous residuals.
```{r echo=FALSE}
#box and whisker plots of specific leaf area by diversity, treatment and species
sla$diversity = as.factor(sla$diversity)
bwplot( ~leaf.sla | diversity, data=sla)
divplot<-ggplot(sla, aes(x=diversity, y=leaf.sla))
divplot + geom_boxplot()
bwplot( ~leaf.sla | taxa, data = sla)
#SOCA, ANGE, and Solidago have the biggest variance. MOFI is pretty tightly grouped, as are HEPA and HEMAX. These should provide good data. ANGE appears to have a significant outlier. Let's find it:
ange.sla<-subset(sla, taxa=="ANGE")
ange.sla$level=as.numeric(ange.sla$level)
ange.sla$leaf.sla=as.numeric(ange.sla$leaf.sla)
angesla<-ggplot(aes(x=paste(ange.sla$block, ange.sla$level), y=leaf.sla), data=ange.sla)
angesla + geom_point()

#C4D1ANGE4 is the major outlier - let's take it out and see if the data becomes closer together:


#write a formula to remove outliers
outlierReplace<-function(dataframe, cols, rows, newValue = NA) {
    if (any(rows)) {
        set(dataframe, rows, cols, newValue)
    }
}

#now remove ouliers from the dataset
my_data<-sla
#View(my_data)
outlierReplace(my_data, "leaf.sla", 
      which(my_data$leaf.sla > 600), NA)

qplot(data = my_data, x = leaf.sla) + ylab("Number of Samples")

write.csv(my_data, file="P:/Research/Thesis/PRP_thesis/Data/sla2.csv")

#from now on use the edited dataset:
sla2<-read.csv("P:/Research/Thesis/PRP_thesis/Data/sla2.csv", header=TRUE)
View(sla2)
```

Create a 'rep' variable combining block, level, and replicate:
```{r}
sla2$rep=paste(sla2$block.num, sla2$level, sla2$replicate)
sla2$rep=as.factor(sla2$rep)
View(sla2)
```

Let's view relationships with the extreme outliers removed:
```{r }
#view relationships between sla and treatment effects
bwplot(leaf.sla ~ treat | factor(diversity), data=sla2)
bwplot(leaf.sla ~ treat + diversity | factor(taxa), data=sla2)
#control actually looks to be lower than all the others, but there are some wicked outliers in the low-diversity plots

#Anovas of differences
treat.aov<-aov(leaf.sla ~ treat + diversity, data=sla2)
s<-summary (treat.aov)
pander(s)
#Treatment is super significant. Why?????
#do a glm to see what the relative effects are of treatments:
treat.glm<-glm(leaf.sla ~ treat*diversity, data=sla2)
summary(treat.glm)
#now nothing is even remotely significant...
#try accounting for some of the variation:
treat.lmer<-lmer(leaf.sla~treat*diversity + (1|block) + (1|level), data=sla2)
summary(treat.lmer)
#nothing has a significant effect escept for the intercept. 

#do a two-way interaction plot:
interaction.plot(sla$treat, sla$diversity, sla$leaf.sla, col = c("red", "blue"))
#perform a two-way mixed-model ANOVA
fit<-lmer(leaf.sla~treat*diversity + (1|rep), data=sla2)
summary(fit)
```

Subset ANGE for analysis:
```{r echo=FALSE}
#subset individual species for analysis
ange.sla<-subset(sla2, taxa=="ANGE")
View(ange.sla)
#ange.sla<-as.data.frame(ange.sla)
ange.aov<-aov(ange.sla$leaf.sla ~ ange.sla$diversity*ange.sla$treat)
summary(ange.aov)

ange.aov2<-aov(leaf.sla~diversity + treat + block + level, data=ange.sla)
summary(ange.aov2)
#Wow. So treatment and diversity were both marginally effective at determining ANGE sla.
###CORRECTED: with the outlier removed, only diversity is significant, but this is pretty highly significant. 
##(p=0.007)

###example script###
#correlation between OM and invasive density
#all invaders
#oplot<-ggplot(invsoil, aes(x=avg.c, y=all.index))
#oplot + geom_point() + geom_smooth()
#o<-oplot + geom_point() + geom_smooth()
###end example script###


#Plot the relationship between ANGE leaf.sla and diveristy using a boxplot
#first make diversity a factor
sla2$diversity = as.factor(sla2$diversity)
ange.sla$diversity = as.factor(ange.sla$diversity)
#now try again:
angeplot<-ggplot(ange.sla, aes(x=diversity, y=leaf.sla), na.rm=TRUE)
angeplot + geom_boxplot()
#Interesting. Many more outliers in the ld category. HD is much more tightly grouped. The >600 outlier was removed for analysis.

####test differences in group means by diversity###
#first, subset data into two groups, ld and hd
ange.sla.hd<-subset(ange.sla, diversity==2)
a.sla.h<-ange.sla.hd[c("leaf.sla")]
ange.sla.ld<-subset(ange.sla, diversity==1)
a.sla.l<-ange.sla.ld[c("leaf.sla")]
#Welch's t-tests for un-transformed data
t.test(a.sla.h, a.sla.l)
#mean of x (hd group) is 176.35, while mean of y (mono) is 156.58. this difference is significant (p=0.00294) 

#compare variances of groups
#var.test(a.sla.h$leaf.sla, a.sla.l$leaf.sla)
#f=0.26538, p=1.682e-8: variances are significantly different from 1. Ratio of variance from sample estimates is 0.2653. 
#qf(0.95, 67, 130)
#f=1.40555, much higher than tabulated f, so there is definitely a difference in group variances.

#try taking the log of both groups to normalize data:
a.sla.h<-log(a.sla.h$leaf.sla)
as.matrix(a.sla.h)
plot(a.sla.h)
a.sla.l<-log(a.sla.l$leaf.sla)
as.matrix(a.sla.l)
plot(a.sla.l)

var.test(a.sla.h, a.sla.l)
qf(0.95, 67, 129)
#variances are closer together but still not overlapping There is too much variation in these groups to really compare their means

#However, looking at their means there is a good argument that they do vary by diversity:
t.test(a.sla.h, a.sla.l)

###plot treatment effects on ANGE sla###
#first get a model for treatment corrected for block and level
angeplot2<-ggplot(ange.sla, aes(x=treat, y=leaf.sla))
angeplot2 + geom_boxplot()                   
#These boxplots did not account for any of the environmental variance.

#try a two-way interaction plot:
ange.interact<-interaction.plot(ange.sla$treat, ange.sla$diversity, ange.sla$leaf.sla, col = c("red", "blue"))
#perform a two-way mixed-model ANOVA
fit<-lmer(leaf.sla~treat*diversity + (1|rep), data=ange.sla)
summary(fit)
#okay, so according to this nothing had a significant influence, or even marginally significant influence. Effects that were significantly correlated with the intercept (ie, no effect, corr >0.5) include diversity, bn, n, rainout, control, and ctrl:diversity.

###EXAMPLE SCRIPT###
#all4<-lmer(all.index ~ (1|rep), data=invsoil, REML=FALSE)
#summary(all4)
#anova(all1.2, all4, test="F")
#anova(all1.2, all4)
###END EXAMPLE###

#I need to create a fitted model for ANGE using block and level as random effects variables:
#do a global model for ANGE:
ange<-lmer(leaf.sla ~ diversity + treat + diversity*treat + (1|rep), data = ange.sla, REML=TRUE)
a<-summary(ange)
a
drop1(ange)
#treatment and diversity are not closely correlated, represent separate effects. 
#All treatments are highly correlated with the intercept, sugesting they may not have a significant effect. 
#All treatments have insignificant correlations. Diversity is the closest to being significant (p=0.145) but still far outsite even the marginal significance range. 
#only 
```


Example from Invader density index script:
ANOVAs for combined fixed and random factors:
```{r}
library(nlme)
library(lmerTest)
###EXAMPLE SCRIPT###
#for cirsium species
#z <- lme(cirsium.index ~ diversity*avg.c, random = ~ 1|rep, data = invsoil) # nlme package
#z <- lmer(cirsium.index ~ diversity*avg.c + (1|rep), data = invsoil) 
#plot(z)
#dat=invsoil
###END EXAMPLE###

#Do lmer for ANGE sla
z<-lmer(leaf.sla ~ diversity + treat + diversity*treat + (1|rep), data=ange.sla)
plot(z)

#Variances are definitely grouped somehow - this is a bit odd.
#try log-transformation:
ange.sla$log.leaf.sla<-log(ange.sla$leaf.sla)
zlog<-lmer(log.leaf.sla ~ diversity*treat + (1|block) + (1|level) + (1|replicate), data=ange.sla)
plot(zlog)
summary(zlog)
#still grouped - this is really weird.

###EXAMPLE SCRIPT###
#model-based fitted means for treatments:
#lsmeans(z, c("diversity", "avg.c"))
#lsmeans(z, c("diversity"))
#az<-anova(z)
#az
#boxplot
#dat=invsoil
#dat$diversity=as.factor(dat$diversity)
#dat$predaz=predict(z)  
#zplot<-ggplot(data=dat, aes(x=diversity, y=predaz, fill=diversity)) + scale_fill_grey(start=0.4, end=1.0) + #geom_boxplot() + labs(x="Diversity", y="Thistle Invasion by diversity*soil OM + BlockID.rand.eff.")
#zplot
###END EXAMPLE SCRIPT###

#Now do it for ANGE:
a<-lsmeans(z, c("diversity", "treat"))
b<-lsmeans(z, c("diversity"))
az<-anova(z)
az
##I don't know why these values are so significant. 

#Compute values for making a boxplot
dat=na.omit(ange.sla)
#View(dat)
dat$predaz=predict(z)
#replacement has 198 rows, data has 199. Need to figure out what to do with NA values?
#used function na.omit to get rid of missing data 

##now plot the predicted values based on the linear mixed-effects model
zplot<-ggplot(data=dat, aes(x=taxa, y=predaz, fill=taxa)) + scale_fill_grey(start=0.4, end=1.0) + geom_boxplot() + labs(x="Diversity", y="SLA(ANGE) by diversity:treatment + random effects")
zplot
#Wow! Groups are different by species!



###EXAMPLE SCRIPT###
###for woody invaders
#z2 <- lme(woody.index ~ diversity*avg.c, random = ~ 1|rep, data = invsoil) # nlme package
#z2 <- lmer(woody.index ~ diversity*avg.c + (1|rep), data = invsoil) 
#plot(z2)
#dat=invsoil
#dat$diversity=as.factor(dat$diversity)
#model-based fitted means for treatments:
#lsmeans(z2, c("diversity", "avg.c"))
#lsmeans(z2, c("diversity"))
#az2<-anova(z2)
#az2

##boxplot
#dat$predaz2=predict(z2)  
#zplot2<-ggplot(data=dat, aes(x=diversity, y=predaz2, fill=diversity)) + scale_fill_grey(start=0.4, #end=1.0) + geom_boxplot() + labs(x="Diversity", y="Woody Invasion by diversity*soil OM + #BlockID.rand.eff.")
#zplot2


###for all invaders
z1 <- lmer(all.index ~ diversity*avg.c + (1|rep), data = invsoil) 
plot(z)
dat=invsoil

#model-based fitted means for treatments:
lsmeans(z, c("diversity", "avg.c"))
lsmeans(z, c("diversity"))
az1<-anova(z1)
az1

#boxplot
dat$diversity=as.factor(dat$diversity)
dat$predaz1=predict(z1)  
zplot1<-ggplot(data=dat, aes(x=diversity, y=predaz1, fill=diversity)) + scale_fill_grey(start=0.4, end=1.0) + geom_boxplot() + labs(x="Diversity", y="Total Invasion by diversity*soil OM + BlockID.rand.eff.")
zplot1


#plot all 3 together:
grid.arrange(zplot1, zplot, zplot2, ncol=3)
```


Subset all Helianthus species for analysis:
```{r}
#subset individual species for analysis
helianthus.sla<-subset(sla2, taxa==c("HEMAX", "HEPA"))
View(helianthus.sla)
helianthus.sla<-as.data.frame(helianthus.sla)
helianthus.aov<-aov(leaf.sla ~ diversity + treat, data=helianthus.sla)
summary(helianthus.aov)
#treatment is significant. Not sure why diversity is not included in the table...
#oh. it's because diversity is not a factor in this table.  
helianthus.sla$diversity<-as.factor(helianthus.sla$diversity)
##let's try using all of the factors:
helianthus.sla2<-aov(leaf.sla~diversity + treat + block + level + diversity, data=helianthus.sla)
summary(helianthus.sla2)
#Treatment and block are significant predictors of Helianthus leaf SLA. The inclusion of site factors block and level make the 

###example script###
#correlation between OM and invasive density
#all invaders
#oplot<-ggplot(invsoil, aes(x=avg.c, y=all.index))
#oplot + geom_point() + geom_smooth()
#o<-oplot + geom_point() + geom_smooth()
###end example script###


#Plot the relationship between ANGE leaf.sla and diveristy using a boxplot
#first make diversity a factor
sla2$diversity = as.factor(sla2$diversity)
ange.sla$diversity = as.factor(ange.sla$diversity)
#now try again:
angeplot<-ggplot(ange.sla, aes(x=diversity, y=leaf.sla), na.rm=TRUE)
angeplot + geom_boxplot()
#Interesting. Many more outliers in the ld category. HD is much more tightly grouped. The >600 outlier was removed for analysis.

####test differences in group means by diversity###
#first, subset data into two groups, ld and hd
ange.sla.hd<-subset(ange.sla, diversity==2)
a.sla.h<-ange.sla.hd[c("leaf.sla")]
ange.sla.ld<-subset(ange.sla, diversity==1)
a.sla.l<-ange.sla.ld[c("leaf.sla")]
#Welch's t-tests for un-transformed data
t.test(a.sla.h, a.sla.l)
#mean of x (hd group) is 176.35, while mean of y (mono) is 156.58. this difference is significant (p=0.00294) 

#compare variances of groups
#var.test(a.sla.h$leaf.sla, a.sla.l$leaf.sla)
#f=0.26538, p=1.682e-8: variances are significantly different from 1. Ratio of variance from sample estimates is 0.2653. 
#qf(0.95, 67, 130)
#f=1.40555, much higher than tabulated f, so there is definitely a difference in group variances.

#try taking the log of both groups to normalize data:
a.sla.h<-log(a.sla.h$leaf.sla)
as.matrix(a.sla.h)
plot(a.sla.h)
a.sla.l<-log(a.sla.l$leaf.sla)
as.matrix(a.sla.l)
plot(a.sla.l)

var.test(a.sla.h, a.sla.l)
qf(0.95, 67, 129)
#variances are closer together but still not overlapping There is too much variation in these groups to really compare their means

#However, looking at their means there is a good argument that they do vary by diversity:
t.test(a.sla.h, a.sla.l)

###plot treatment effects on ANGE sla###
#first get a model for treatment corrected for block and level
angeplot2<-ggplot(ange.sla, aes(x=treat, y=leaf.sla))
angeplot2 + geom_boxplot()                   
#These boxplots did not account for any of the environmental variance.
#I need to create a fitted model for ANGE using block and level as random effects variables:

###EXAMPLE SCRIPT###
#all4<-lmer(all.index ~ (1|rep), data=invsoil, REML=FALSE)
#summary(all4)
#anova(all1.2, all4, test="F")
#anova(all1.2, all4)
###END EXAMPLE###

#do a global model for ANGE:
ange<-lmer(leaf.sla ~ diversity *treat + (1|block) + (1|level) + (1|replicate), data = ange.sla, REML=FALSE)
a<-summary(ange)
a



```

